---
title: "Predicting Performance in the NBA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = FALSE}
#install.packages("readxl")
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
# install.packages("caret",
#                  repos = "http://cran.r-project.org",
#                  dependencies = c("Depends","Imports","Suggests"))
library(caret)
#install.packages("corrplot")
library(corrplot)
#install.packages("e1071")
library(e1071)
#install.packages("RANN")
library(RANN)
#install.packages("randomForest")
library(randomForest)
#install.packages("skimr")
library(skimr)
#install.packages("dplyr")
library(dplyr)
#install.packages("plotly")
library(plotly)
#install.packages("DT")
#install.packages("htmlwidgets")
library(DT)
```

```{r data, include = FALSE}
source("Scripts/Wrangling.R")
source("Scripts/Preprocessing.R")
```

Unbalanced Panel of `r length(unique(df$Player))` NBA players from `r min(df$Year_t3, na.rm = TRUE)` to `r max(df$Year_t1, na.rm = TRUE)+1`: Outcome Value over replacemente player (VORP) and `r length(df)-1` features before preprocessing.

```{r correlations}
#create correlation matrix
cor_num <- cor(df_num_mean)

#delete suggested variables with high correlation with other features
cor_high <- findCorrelation(cor_num, cutoff = .95)  

names(df_num_mean[,cor_high]) # deleted variables due to collinearity

```

```{r show data}
skim(x_mean_train)
```


```{r modeling, cache = TRUE}
# Hyperparameter Grid for Random Forest
rfGrid <- expand.grid(
    mtry = seq(round(length(x_mean_train)/3-5,0), round(length(x_mean_train)/3+5,0), by=2),
    #mtry = round(length(x_mean_train)/3,0),
    splitrule = "variance",
    min.node.size = 5
  )

# Random Forest via Caret's Train function & ranger lib
set.seed(1234)
rf <- train(x = x_mean_train, y = y_train,
            method = "ranger",
            num.trees = 500,
            tuneGrid = rfGrid,
            trControl = trainControl(method = "cv"),
            importance = "permutation")
```

```{r modeling results}
plot(rf)
```


```{r pred}

#predict yhat for test set
rfPred <- predict(rf, x_mean_test) 

#data frame to compare predictions to outcomes in test data
predictions <- data.frame(Player = df_pred$Player,
                          Year = df_pred$Year_t1+1,
                          Pos = df_pred$Pos_t1,
                          y = y_test,
                          y_hat = rfPred) %>% 
  mutate(error = y_hat - y) %>% 
  mutate_if(is.factor, as.character)

#predictions df for training data
predictions_training <- data.frame(Player = df[trainingRow, ]$Player,
                                   Year = df[trainingRow, ]$Year_t1+1,
                                   Pos = df[trainingRow, ]$Pos_t1,
                                   y = y_train,
                                   y_hat = predict(rf)) %>% 
  mutate(error = y_hat - y)
```

## Predictions Player Performance {.tabset .tabset-fade .tabset-pills}

### Test data

The test data has `r length(predictions$Player)` observations with mean y of `r round(mean(predictions$y),2)` and standard deviation of `r round(sd(predictions$y),2)`. The RMSE is `r sqrt(mean(predictions$error^2))`; the MAE is `r mean(abs(predictions$error))`.

```{r test data, echo = FALSE, warning = FALSE}

predictions$y_hat <- round(predictions$y_hat,2)

#ggplot visual
p <- ggplot(predictions, aes(x=y_hat, y = y, color = Pos,
                             text = paste0(Player, " ", Year))) +
  geom_jitter(alpha=0.4, size = 1)+
  geom_abline(linetype = 2)+
  scale_x_continuous(breaks = seq(0,15,1))+
  scale_y_continuous(breaks = seq(0,15,1))+
  theme_bw()

#plotly wrapper for interactivity
ggplotly(p)
```

### Training data

The test data has `r length(predictions_training$Player)` observations with mean y of `r round(mean(predictions_training$y),2)` and standard deviation of `r round(sd(predictions_training$y),2)`.

```{r training data, echo = FALSE, warning = FALSE}

predictions_training$y_hat <- round(predictions_training$y_hat,2)

q <- ggplot(predictions_training, aes(x=y_hat, y = y, color = Pos,
                             text = paste0(Player, " ", Year))) +
  geom_jitter(alpha=0.4, size = 1)+
  geom_abline(linetype = 2)+
  scale_x_continuous(breaks = seq(0,15,1))+
  scale_y_continuous(breaks = seq(0,15,1))+
  theme_bw()


ggplotly(q)
```

```{r testing 2018, include = FALSE, cache = TRUE}
# source("Scripts/Preprocessing_test2018.R")
# 
# source("Scripts/RandomForest.R")
# 
# source("Scripts/Rookies.R")
# 
# source("Scripts/Team_Records.R")
# 
# predictions_2018 <- data.frame(Player = df_pred$Player,
#                           Year = df_pred$Year_t1+1,
#                           Pos = df_pred$Pos_t1,
#                           y = y_test,
#                           y_rf_mean = rfPred) %>% 
#   mutate(error_rf_mean = y_rf_mean - y) %>% 
#   mutate_if(is.factor, as.character)


```

```{r test data 2018, echo = FALSE, warning = FALSE}

# predictions_2018$y_rf_mean <- round(predictions_2018$y_rf_mean,2)
# 
# p <- ggplot(predictions_2018, aes(x=y_rf_mean, y = y, color = Pos,
#                              text = paste0(Player, " ", Year))) +
#   geom_jitter()+
#   geom_abline(linetype = 2)+
#   scale_x_continuous(breaks = seq(0,15,1))+
#   scale_y_continuous(breaks = seq(0,15,1))+
#   theme_bw()


#ggplotly(p)
```

```{r teams 2018, include = FALSE}
# adv2018 <- read_excel("Data/2018_adv.xlsx") %>%
#   subset(Tm != "TOT") %>%
#   group_by(Player) %>%
#   top_n(n=1,G)
# 
# df2018 <- merge(predictions_2018, adv2018, by = "Player") %>%
#   dplyr::select(Player, Team = Tm, G, MP, vorp = VORP, prediction_vorp = y_rf_mean)
# 
# rookies_2018 <- df_rookies %>%
#   filter(Year == 2018)
# 
# rookies_2018 <- merge(rookies_2018, adv2018, by = "Player") %>%
#   dplyr::select(Player, Team = Tm.y, G, MP, vorp = VORP, prediction_vorp = vorp_pred)
# 
# df2018 <- rbind(df2018,rookies_2018)

# df_2018_Tm <- df2018 %>% 
#   group_by(Team) %>% 
#   dplyr::summarize(vorp = sum(prediction_vorp)) 
# 
# df_2018_Tm$wins_pred = predict(lm_teams, newdata = df_2018_Tm)
# 
# teams2018 <- subset(teams_all, teams_all$Year == 2018) %>% 
#   dplyr::select(Team, Wins, VORP = vorp)
# 
# df_2018_Tm <- merge(df_2018_Tm, teams2018, by = "Team")
```

```{r team performance 2018, echo = FALSE, warning = FALSE}
# df_2018_Tm$Wins_pred = round(df_2018_Tm$wins_pred,0)
# 
# q <- ggplot(df_2018_Tm, aes(x=Wins_pred, y = Wins, text = Team))+
#   geom_point()+
#   # geom_text(aes(label = Team), vjust = -1)+
#   geom_abline(linetype = 2)+
#   scale_x_continuous(breaks = seq(0,70,5))+
#   scale_y_continuous(breaks = seq(0,70,5))+
#   theme_bw()
# 
# ggplotly(q)
```

## 2018/19 season {.tabset .tabset-fade .tabset-pills}

```{r 2019 predictions, include = FALSE, cache = TRUE}
source("Scripts/2019/2019_Main.R")
```


### Teams

```{r}

predictions_2019 %>% 
  mutate(vorp = round(vorp,2),
         Wins_pred_2019 = round(Wins_pred_2019,2),
         Error_Pred_Vegas = round(Error_Pred_Vegas,2),
         Error_Pelton_Vegas = round(Error_Pelton_Vegas,2),
         Sum_Error = round(Sum_Error,2)
         ) %>% 
datatable(options = list(scrollX = TRUE))
```

```{r team performance 2019, echo = FALSE, warning = FALSE, cache = TRUE}
q <- ggplot(predictions_2019, aes(y=Wins_pred_2019, x = Wins_Vegas_2019, text = Team))+
  geom_point(aes(color = Sum_Error, size = abs(Sum_Error)))+
  # geom_text(aes(label = Team), vjust = -1)+
  geom_abline(linetype = 2)+
  scale_x_continuous(breaks = seq(0,70,5))+
  scale_y_continuous(breaks = seq(0,70,5))+
  scale_color_gradient(low = "red", high = "limegreen")+
  theme_bw()

ggplotly(q)
```

### Player growth
```{r player performance 2019, echo = FALSE, warning = FALSE, cache = TRUE}
p <- ggplot(predictions2019, aes(x = vorp_t1, y = vorp_pred, text = Player)) +
  geom_jitter(aes(color = Tm))+
  # geom_text(aes(label = Team), vjust = -1)+
  geom_abline(linetype = 2)+
  theme_bw()

ggplotly(p)
```